The impact of gas bulk rotation on the Lyman-$\alpha$ line

### 2013-08-16 paper/Analisis.tex
number of scatterings

To quantify the line broadening we measure a modified version of the full
width at half maximum (FWHM). Instead we measure it only for half of
the line, $W_{1/2}$. This definition allows us to quantify the line
width both in the cases of double and single peak emission.  In the
case of double peaked emission $W_{1/2}$ corresponds to the width of a
single peak, while in the extreme case of high rotational velocities,
when the double peak is erased, it simply correspond to half FWHM.

@@

Figure \ref{fig:maximumsvsvelocity} shows the position for the peak
maxima as a function of the rotational velocity $V_{\rm max}$. This
figure shows us clearly that in the case of central distributed
sources there is barely any change with rotational velocity in the
range of explored parameters. However, in the case of
homogeneously emitted sources the maxima position remain close to
constant until beyond some velocity threshold the line becomes single
peaked with $x_{\rm max}=0$.

It is probable that the transition to a single peak line occurs for systems
where it  becomes easier for the bulk of the photons to escape with the lowest
number of scatterings possible, allowing them not to move very far
from the center of the line. This hypothesis can help to explain how the
single peak stage is easily achieved in the homogeneous source
distribution. In this case there is a fraction of the photons that are
inside a photosphere region with $\tau_{H,r}\ll \tau_{H}$ where
$\tau_{H,r}$ is the optical depth from the radius of emission to the
sphere's surface. This conditions allows the photons to escape with
much less scatterings compared to the photons emitted at the very
center of the sphere. In turn, it gives the photons less scatterings
to be placed far from the line center. Increasing the rotational
velocity $V_{\rm max}$ reduces the optical depth making the
photosphere region effectively larger, increasing the number of
photons escaping close to the lines's center. For the central case the
photosphere is not present, and other mechanisms must be at play in
the steady reduction of the double peak.

For larger values of $\tau_{H}$ the transitional velocity $V_{\rm
  trans}$ is also larger. In our models we find the following
correspondence between the optical depth $\tau_{\rm H}=\{10^5, 10^6,
10^7\}$ and the transitional velocities $V_{\rm
  trans}=\{50-100\kms, 100-200\kms, 200-300\kms\}$ which can only be
constrained to be in the range of velocities in the models that gave a $x_{\rm
  max}\neq 0$ and $x_{\rm max}=0$.   In the next section explore the
origin of this trends and consider to what extent the hypothesis of a low
number of scatterings is correlated with the emergence of a single
peak.

@@

    \caption{2D histogram of $N_{scatt}$ vs $x$. The upper (lower) pannels
      show the homogeneous (central) source distribution. Left
      corresponds to the static case and the right
      $V_{max}=300km/s$. The colour scale is locarithmic\label{fig:Nscatt2D}}  
\end{figure*}

The number of times that a \ly photon is absorbed and re-emitted is
connected to the final frequency that it can have after escaping the
galaxy. In the case of static gas geometries, a large value of the
optical depth is immediately followed by a high number of
scatterings. In turn a large optical depth increases the probability
that a \ly photon will be found far from the center of the line. The
peak maxima shifting away form the line center as the amount of
neutral hydrogen increases.

In Figure \ref{fig:Nscatt} we see the average number of scatterings
$\langle N_{\rm scatt}\rangle$ as a function of the rotational velocity
$V_{\rm max}$. For the central distributions we find that there is not
a significant change for increasing rotational
velocities, $\langle N_{\rm scatt}\rangle$ changes less than $0.5\%$
for different velocities. In this case we also find that the average
number of scatterings is proportional to the optical depth, as
expected in analogy from the analytic result for the homogeneous
infinite-slab $\langle N_{\rm  scatt}\rangle=1.612\tau_{\rm   H}$
\citep{Adams72,Harrington73}. In our experiments we find  $\langle N_{\rm
  scatt}\rangle= (1.505, 1.002, 0.918)\tau_{\rm   H}$, for $\tau_{\rm H} = (10^{5}, 10^{6}, 10^{7})$ respectively.

However, for the homogeneous distribution there is a clear decrease of
$\langle N_{\rm  scatt}\rangle$ as the  $V_{\rm max}$ increases. This
effect more pronounced for the lower values of the optical depth. For
$\tau_{\rm H}=10^5$ the average number of scatterings decreases by
$61\%$ at $V_{\rm max}=300\kms$ in comparison to the static case.

The analytic expectation for the slab with homogeneously emitted
sources is $\langle N_{\rm  scatt}\rangle=1.16\tau_{\rm   H}$
\citep{Harrington73}, a factor of $0.72$ lower than the case of the
centrally emitted photons. In our case we find that for the static
spheres this factor is $(1.510553, 1.002823, 0.9188793)$ for $\tau_{\rm H} = (10^{5}, 10^{6}, 10^{7})$ respectively.

>>>>

To quantify the line broadening we use a modified version of the full
width at half maximum (FWHM). We measure it only for half of
the line, $W_{1/2}$. This definition allows us to quantify the line
width and see the transition from double to single peak emission.  In the
case of double peaked emission $W_{1/2}$ corresponds to the width of a
single peak, while in the extreme case of high rotational velocities,
when the double peak is erased, it simply correspond to half FWHM.

@@

Figure \ref{fig:maximumsvsvelocity} shows the position for the peak
maxima as a function of the rotational velocity $V_{\rm max}$. This
Figure shows us clearly that in the case of central distributed
sources there is barely any change with rotational velocity in the
range of explored parameters. However, in the case of
homogeneously emitted sources the maxima position remain close to
constant until beyond some velocity threshold the line becomes single
peaked with $x_{\rm max}=0$.

The transition to a single peak line occurs for systems
where it  becomes easier for the bulk of the photons to escape with the lowest
number of scatterings possible, allowing them not to move very far
from the center of the line. This explains how the single peak stage
is easily achieved in the homogeneous source distribution. In this
case there is a fraction of the photons that are inside a photosphere
region with $\tau_{H,r}\ll \tau_{H}$ where $\tau_{H,r}$ is the optical
depth from the radius of emission to the sphere's surface. This
conditions allows the photons to escape with much less scatterings
compared to the photons emitted at the very center of the sphere. In
turn, it gives the photons less scatterings to be placed far from the
line center. Increasing the rotational velocity $V_{\rm max}$ reduces
the optical depth making the photosphere region effectively larger,
increasing the number of photons escaping close to the lines's
center. For the central case the photosphere is not present, and other
mechanisms must be at play in the steady reduction of the double peak.

For larger values of $\tau_{H}$ the transition to a single peaked
emission occurs for larger values of the rotation velocity. In our
models we find the following correspondence between the optical depth
$\tau_{\rm H}=\{10^5, 10^6, 10^7\}$ and the transitional velocities $V_{\rm
  trans}=\{50-100\kms, 100-200\kms, 200-300\kms\}$ which can only be constrained to be in the range of velocities in the models that gave a $x_{\rm
  max}\neq 0$ and $x_{\rm max}=0$.   In the next section explore the
origin of this trends and show to what extent a low number of
scatterings is correlated with the emergence of a single peak.

@@

    \caption{2D histogram of $N_{\rm scatt}$ vs $x$. The upper (lower) pannels
      show the homogeneous (central) source distribution. Left
      corresponds to the static case and the right
      $V_{max}=300km/s$. The colour scale is logarithmic on the
      number of photons with given values of $N_{\rm scatt}$ and
      $x$. \label{fig:Nscatt2D}}   
\end{figure*}

The number of times that a \ly photon is absorbed and re-emitted is
connected to the final frequency that it can have after escaping the
galaxy. In the case of static gas geometries, a large value of the
optical depth is immediately followed by a high number of
scatterings. In turn a large optical depth increases the probability
that a \ly photon to be found far from the center of the line. In this
case, the peak maxima shift away form the line center as the amount of
neutral hydrogen increases.

In Figure \ref{fig:Nscatt} we show the average number of scatterings
$\langle N_{\rm scatt}\rangle$ as a function of the rotational velocity
$V_{\rm max}$. For the central distributions we find that there is not
a significant change for increasing rotational
velocities, $\langle N_{\rm scatt}\rangle$ changes less than $0.5\%$
for different velocities. In this case we also find that the average
number of scatterings is proportional to the optical depth, as
expected in analogy from the analytic result for the homogeneous
infinite-slab $\langle N_{\rm  scatt}\rangle=1.612\tau_{\rm   H}$
\citep{Adams72,Harrington73}. In our experiments we find  that for the
static spheres with centraly distributed sources $\langle N_{\rm
  scatt}\rangle= (1.50, 1.00, 0.92)\tau_{\rm   H}$. for optical depths
$\tau_{\rm H} = (10^{5}, 10^{6}, 10^{7})$

Figure \ref{fig:Nscatt} shows that for the homogeneous distribution
there is a clear decrease of $\langle N_{\rm  scatt}\rangle$ as the
$V_{\rm max}$ increases. This effect more pronounced for the lower
values of the optical depth. For $\tau_{\rm H}=10^5$ the average
number of scatterings decreases by $61\%$ at $V_{\rm max}=300\kms$ in
comparison to the static case.

The analytic expectation for the slab with homogeneously emitted
sources is $\langle N_{\rm  scatt}\rangle=1.16\tau_{\rm   H}$
\citep{Harrington73}, a factor of $0.72$ lower than the case of the
centrally emitted photons. In our case we find that for the static
, while for homogeneously distributed source $\langle N_{\rm
  scatt}\rangle= (1.50, 1.00, 0.92)\tau_{\rm   H}$. {\bf seguro ????}
